---
title: "ingest-data"
author: "JJayes"
date: "29/09/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(readxl)
library(scales)
# library(sf)
```


## Purpose

Ingest data from regional accounts 

```{r}
df <- read_excel("data/regional-accounts/import.xlsx")

df %>% 
    pivot_longer(-`GDP`) %>% 
    relocate(name, .before = `GDP`) %>% 
    rename(year = name) %>% write.csv("data/regional-accounts/exp.csv")

```

## Figures

Import

```{r}
df <- read_excel("data/regional-accounts/regional-accounts-aggregated.xlsx")
```

### What figures?

```{r}
df %>% 
    distinct(series)
```

### Labour force in different sectors

```{r}
df %>% 
    filter(series == "Labour force shares in the Swedish counties",
           lan == "Stockholms län") %>% 
    ggplot(aes(year, value, fill = subset)) +
    geom_col()
```

```{r}
df %>% 
    filter(series == "Labour force shares in the Swedish counties",
           subset == "Agriculture") %>% 
    ggplot(aes(year, value, color = lan)) +
    geom_line() +
    scale_y_continuous(labels = percent_format()) + 
    scale_color_viridis_d() +
    theme(legend.position = "none") +
    labs(y = "Share of labour force in Agriculture",
         x = NULL)

```

```{r}
library(ggiraph)

gg <- df %>%
  filter(
    series == "Labour force shares in the Swedish counties",
    subset == "Mining"
  ) %>%
  ggplot(aes(year, value, color = lan)) +
  geom_line_interactive(aes(tooltip = lan, data_id = lan)) +
  scale_y_continuous(labels = percent_format()) +
  scale_color_viridis_d() +
  theme(legend.position = "none") +
  labs(
    y = "Share of labour force in Agriculture",
    x = NULL
  )


x <- girafe(ggobj = gg, width_svg = 8, height_svg = 6,
  options = list(
    opts_hover_inv(css = "opacity:0.1;"),
    opts_hover(css = "stroke-width:2;")
  ))
x
```

### Shares of total labour force by county

maybe geom ridges??

```{r}
df %>% 
    filter(series == "Shares of total labour force by county") %>%
    ggplot(aes(year, value, fill = lan)) +
    geom_area()
```


### Wages in the Swedish counties (1750-1850, Sweden=1)

```{r}
df %>% 
    filter(series == "Wages in the Swedish counties (1750-1850, Sweden=1)") %>%
    ggplot(aes(year, value, colour = lan)) +
    geom_line() +
    facet_wrap(~ subset)
```

### GDP per capita in the Swedish counties (1750-1850, Sweden=100)

```{r}
df %>% 
    filter(series == "GDP per capita in the Swedish counties (1750-1850, Sweden=100)") %>%
    ggplot(aes(year, value, colour = lan)) +
    geom_line() +
    facet_wrap(~ subset)

gg <- df %>%
  filter(series == "GDP per capita in the Swedish counties (1750-1850, Sweden=100)") %>%
  ggplot(aes(year, value, color = lan)) +
  geom_line_interactive(aes(tooltip = lan, data_id = lan)) +
  scale_y_continuous(labels = number_format()) +
  scale_color_viridis_d() +
  theme(legend.position = "none") +
  labs(
    y = "GDP per capita (Sweden 1750 = 100)",
    x = NULL
  )


x <- girafe(
  ggobj = gg, width_svg = 8, height_svg = 6,
  options = list(
    opts_hover_inv(css = "opacity:0.1;"),
    opts_hover(css = "stroke-width:2;")
  )
)
x
```

### Shares of total GDP in the Swedish counties

```{r}
lan_map <- st_read("data/county-shapefiles/maps/sverige-lan-counties-of-sweden.shp") %>% 
    rename(lan = lan_namn)

s_map <- lan_map %>%
  mutate( # JavaScript call to open website 
    onclick = glue::glue(
'window.open("https://en.wikipedia.org/wiki/{lan}")')
  ) %>%
  ggplot() +
  geom_sf_interactive(
    aes(data_id = lan, 
        tooltip = lan,
        onclick = onclick)
  ) +
  theme_void()

# girafe(
#   ggobj = s_map,
#   width_svg = 6,
#   height_svg = 6*0.618
# )
```


```{r}
library(cowplot)
int_fig <- girafe(
  ggobj = plot_grid(gg, s_map, rel_widths = c(1.2, 1), ncol = 2),
  width_svg = 5,
  height_svg = 3,
  options = list(
    # opts_tooltip(css = tooltip_css, delay_mouseover = 0, delay_mouseout = 0),
    opts_hover_inv(css = "opacity:0.1;"),
    opts_hover(css = "stroke-width:2;")  )
)

int_fig
```

### Shares of value added by sector in the Swedish counties, (1750-1850)

```{r}
df %>% 
    filter(series == "Shares of value added by sector in the Swedish counties, (1750-1850)",
           year != 1769) %>% 
    select(lan, year, subset, value) %>% 
    pivot_wider(names_from = year, values_from = value, values_fill = 0) %>% 
    pivot_longer(-c(lan, subset), names_to = "year") %>% 
    mutate(year = parse_number(year)) %>% 
    ggplot(aes(year, value, fill = subset)) +
    geom_area() +
    facet_wrap(~ lan)

```


```{r}
df %>%
  filter(
    series == "Shares of value added by sector in the Swedish counties, (1750-1850)",
    year != 1769,
    year == 1800
  ) %>%
  select(lan, year, subset, value) %>%
  pivot_wider(names_from = year, values_from = value, values_fill = 0) %>%
  pivot_longer(-c(lan, subset), names_to = "year") %>%
  inner_join(lan_map) %>%
  ggplot() +
  geom_sf(aes(geometry = geometry, fill = value)) +
  facet_wrap(~subset) +
  scale_fill_gradient2_interactive(
    low = "blue",
    high = "red",
    mid = "pink",
    midpoint = 0.4,
    labels = percent_format()
  ) +
  theme_void() +
  theme(legend.position = "bottom") +
  guides(fill = guide_legend(label.position = "top", label.hjust = 1)) +
  labs(fill = "Share of value added by sector")
```

## Maps

Read in kommun shapefiles

```{r}
shp <- sf::st_read("data/kommun shapefiles/Kommun_Sweref99TM_region.shp")

shp %>% ggplot() +
  geom_sf()

```

```{r}

library(glue)

year_filter <- 1850

s_map_2 <- lan_map %>%
  inner_join(df %>% filter(
    series == "Shares of total GDP in the Swedish counties (1750-1850)",
    year == year_filter
  )) %>%
  mutate( # JavaScript call to open website
    onclick = glue::glue(
      'window.open("https://en.wikipedia.org/wiki/{lan}")'
    )) %>%
  ggplot() +
  geom_sf_interactive(
    aes(
      fill = value,
      data_id = lan,
      tooltip = lan,
      onclick = onclick
    )
  ) +
  scale_fill_gradient2_interactive(
    low = "blue",
    high = "red",
    mid = "pink",
    midpoint = 0.05,
    labels = percent_format()
  ) +
  labs(fill = glue("Share of GDP in {year_filter}")) +
  theme_void()


girafe(
  ggobj = s_map_2,
  width_svg = 6,
  height_svg = 6*0.618
)
```


```{r}
chart_col <- df %>%
  filter(
    series == "Shares of total GDP in the Swedish counties (1750-1850)",
    year == year_filter
  ) %>%
  mutate(lan = fct_reorder(lan, value)) %>%
  ggplot(aes(value, lan, fill = value)) +
  geom_col_interactive(aes(data_id = lan, tooltip = lan), show.legend = F) +
  scale_x_continuous(labels = percent_format()) +
  scale_fill_gradient2_interactive(
    low = "blue",
    high = "red",
    mid = "pink",
    midpoint = 0.05,
    labels = percent_format()
  ) +
  labs(
    x = NULL,
    y = NULL
  )

girafe(
  ggobj = chart_col,
  width_svg = 6,
  height_svg = 6*0.618
)
```

```{r}
int_fig_2 <- girafe(
  ggobj = plot_grid(chart_col, s_map_2, rel_widths = c(4, 1), ncol = 2),
  width_svg = 5,
  height_svg = 3,
  options = list(
    # opts_tooltip(css = tooltip_css, delay_mouseover = 0, delay_mouseout = 0),
    opts_hover_inv(css = "opacity:0.1;"),
    opts_hover(css = "stroke-width:2;")  )
)

int_fig_2
```


### Population and GDP over time

```{r}
chart_pop_gdp <- df %>% 
    filter(series %in% c("Population in the Swedish counties", "GDP per capita in the Swedish counties (1750-1850, Sweden=100)")) %>% 
    select(-subset, -source) %>% 
    mutate(series = case_when(
         series == "GDP per capita in the Swedish counties (1750-1850, Sweden=100)" ~ "GDP per capita (Sweden 1750 = 100)",
         series == "Population in the Swedish counties" ~ "Population"
    )) %>% 
    ggplot(aes(year, value, colour = lan)) +
    geom_line_interactive(aes(tooltip = lan, data_id = lan), show.legend = F) +
    facet_wrap(~series, scales = "free_y", nrow = 2) +
    scale_y_continuous(labels = number_format()) +
    scale_colour_viridis_d_interactive() +
    labs(x = NULL,
         y = NULL)

girafe(
  ggobj = chart_pop_gdp,
  width_svg = 6,
  height_svg = 6 * 0.618,
  options = list(
    # opts_tooltip(css = tooltip_css, delay_mouseover = 0, delay_mouseout = 0),
    opts_hover_inv(css = "opacity:0.1;"),
    opts_hover(css = "stroke-width:2;")
  )
)
```

```{r}
df %>% 
    filter(series %in% c("Population in the Swedish counties", "GDP per capita in the Swedish counties (1750-1850, Sweden=100)", "Shares of total GDP in the Swedish counties (1750-1850)")) %>% 
    select(-subset, -source) %>% 
    mutate(series = case_when(
         series == "GDP per capita in the Swedish counties (1750-1850, Sweden=100)" ~ "GDP per capita",
         series == "Population in the Swedish counties" ~ "Population",
         series == "Shares of total GDP in the Swedish counties (1750-1850)" ~ "Share of GDP"
    )) %>% 
    pivot_wider(names_from = series, values_from = value) %>%
    mutate(`Share of GDP` = percent(`Share of GDP`, accuracy = .1),
           Population = format(Population, big.mark = " ")) %>% 
    arrange(year) %>% 
    DT::datatable()
```


## New shapefile-map

```{r}
shp <- st_read("data/kartfiler/ln74_96bj_region.shp")
lan_names <- read_excel("data/lan_names.xlsx")

shp <- shp %>% 
    mutate(index = row_number()) %>% 
    inner_join(lan_names, by = "index")

shp %>% 
    ggplot() +
    geom_sf()

shp %>% write_rds("data/kartfiler/Sweden-24-counties.rds")

```

